---
title: "Practice exam"
output: html_document
date: "2025-03-03"
---
#Load the data
```{r}
library(fpp2)
library(ggplot2)
data("arrivals")
us_arrivals <- arrivals[,"US"] 
us_arrivals <- ts(arrivals[, "US"], start=start(arrivals), frequency=frequency(arrivals))

```
#Time series plot
```{r}
autoplot(us_arrivals) + ggtitle("US Arrivals to Australia") +  ylab("Number of Arrivals") + xlab("Year")
```
###time series apears to have an upward trend, seasonal fluctuation, periodic peaks

###2 - Central tendency
#summarize the statistics
```{r}
summary(us_arrivals) # Min, Max, Mean, Median, Quartiles
boxplot(us_arrivals, main="Boxplot of US Arrivals to Australia", col="lightblue")
```
###Observations from summary stats and boxplot
-median and mean provide central tendence
-interquartile range shows variability in arrvials
-may be some outlier 



#3- Decomposition
```{r}
dec <- decompose(us_arrivals, type = "multiplicative") 
autoplot(dec)
```
###
-time series shows seasonality
-decomposition is multiplicative since it increases over time
-it is seasnoal peakign in december, most likely due to holiday season travels
-Value low for february and high in december
December may be Christmas/New Year season and February being an off season
#seasonaly adjusted series
```{r}
seasonally_adjusted <- us_arrivals / dec$seasonal

autoplot(us_arrivals) + autolayer(seasonally_adjusted, series="Seasonally Adjusted", color="red") +ggtitle("US Arrivals: Seasonally Adjusted vs Original") +ylab("Number of Arrivals") +xlab("Year")
```
#seasonality shows significant fluctuations 



#3- Naive Method
```{r}
naive_model <- naive(us_arrivals, h=24) 
autoplot(naive_model)
```
#Residual analysis
```{r}
residuals_naive <- residuals(naive_model)
autoplot(residuals_naive) + ggtitle("Residuals of Naïve Model")

```
#Hisogram of residuals 
```{r}
hist(residuals_naive, main="Histogram of Residuals", xlab="Residuals", col="lightblue")
```
#fitted values vs residuals
```{r}
plot(fitted(naive_model), residuals_naive, main="Fitted Values vs Residuals", xlab="Fitted Values", ylab="Residuals")
abline(h=0, col="red")
```
#Actual values vs residuals
```{r}
plot(us_arrivals, residuals_naive, main="Actual Values vs Residuals", xlab="Actual Values", ylab="Residuals")
abline(h=0, col="red")
```
#Acf of residuals
```{r}
Acf(residuals_naive, main="ACF of Residuals")
```

###
-the residuals should be randomly distributed around zero
-the histogram shows near normal distribution
-the acf plot indicated autocorelation


#forecasting
```{r}
fc <- forecast(naive_model, h=24) 
autoplot(fc) +
  ggtitle("Forecast for Next Year") +
  ylab("Number of Arrivals") +
  xlab("Year")
```
#Forecasting table 
```{r}
as.data.frame(fc)
```
#forecasting accuracy

```{r}
accuracy(naive_model)
```


###
the forecast shows a continuation past trends with seasonalfluctuations, 
